NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-05-22 Manuscript/manuscriptPS.tex
Update manuscript and some cleaning

%In NMRlipids I and II project we were looking for a MD model
%which would correctly reproduce headgroup and glycerol
%backbone structures and cation binding for PC lipid bilayers \cite{botan15,catte16}.
%Here we extend the same goal for lipids with negatively charged PS headgroup.
%Chemical structure of PS headgroup together with other common biological
%lipids is shown in Fig. \ref{lipids}.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/lipids.pdf}
  \caption{\label{lipids}
    Chemical structures and labels for the headgroup carbons.
  }
\end{figure}

@@

\subsection{Solid state NMR experiments}
The experimental protocol is the same used in Ref. \citenum{ferreira16}.
\todo{Some basic details should be given.}

@@

The headgroup and glycerol backbone order parameters of 
POPS measured in this work are compared to the literature values of
DOPS \cite{browning80} and POPC \cite{ferreira13} in Fig. \ref{HGorderParameters}. 
Our results for POPS are in good agreement with the previously reported
values for DOPS measured with $^2$H NMR. Significant differences are
observed between PC and PS lipids, especially at the headgroup region.  
Previous discussions in the literature have concluded that  
glycerol backbone structure is largely similar in
PC, PE, PG and PS lipids \cite{gally81}. The headgroup region was
found to be similar in PC, PE and PG lipid (assuming that the signs of PE and PG
order parameters are the same as in PC), while the PS headgroup was suggested
to be more rigid \cite{wohlgemuth80,buldt81}. 
The detailed structural differences between the headgroups is, however, not known.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPCPS.eps}
  \caption{\label{HGorderParameters}
    Headgroup and glycerol backbone order parameters of POPS measured in this work compared
    with values for DOPS ($^2$H NMR, 0.1M of NaCl) \cite{browning80} and 
    POPC  ($^{13}$C NMR) \cite{ferreira13} from literature. Signs for PS order parameters
    as measured in this work and signs for PC as measured in Refs \cite{??,ferreira16}.
    %POPG from \cite{borle85} contains 10nM PIPES, DPPG from \cite{wohlgemuth80} contains  10mM PIPES and 100mM NaCl,
    %DPPE from \cite{seelig76}, E.coliPE and E.coliPG are from \cite{gally81}.
  }
\end{figure}

>>>>

%In NMRlipids I and II project we were looking for a MD model
%which would correctly reproduce headgroup and glycerol
%backbone structures and cation binding for PC lipid bilayers \cite{botan15,catte16}.
%Here we extend the same goal for lipids with negatively charged PS headgroup.
%Chemical structure of PS headgroup together with other common biological
%lipids is shown in Fig. \ref{lipids}.
%\begin{figure}[]

%  \centering
%  \includegraphics[width=9.0cm]{../Figs/lipids.pdf}
%  \caption{\label{lipids}
%    Chemical structures and labels for the headgroup carbons.
%  }
%\end{figure}

@@

\subsection{Solid state NMR experiments}
The magnitude and signs of the C-H bond order parameters in
headgroup and glycerol backbone were measured
using natural abundance $^{13}$C solid state NMR spectroscopy as
described previously \cite{ferreira13,ferreira16}.
Shortly, the absolute values of the order parameters were determined from the dipolar splittings
given by the indirect dimension of 2D R-PDFL experiment \cite{dvinskikh04} and
the signs were measured using S-DROSS experiments \cite{gross97}.

\todo{Details of the used spectrometer and maybe some other details should be given.}
\todo{Sample preparation should be described.}

@@

The headgroup and glycerol backbone order parameters of 
POPS measured in this work are compared to the literature values of
DOPS \cite{browning80} and POPC \cite{ferreira13} in Fig. \ref{HGorderParameters}. 
Our results for POPS are in good agreement with the previously reported
values for DOPS measured with $^2$H NMR. Significant differences are
observed between PC and PS lipids, especially at the headgroup region.  
Previous discussions in the literature have concluded that  
glycerol backbone structure is largely similar in
PC, PE, PG and PS lipids \cite{gally81}. The headgroup region was
found to be similar in PC, PE and PG lipid (assuming that the signs of PE and PG
order parameters are the same as in PC), while the PS headgroup was suggested
to be more rigid \cite{wohlgemuth80,buldt81}. 
The detailed structural differences between the headgroups is, however, not known.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/PCPScomp.pdf}
  \caption{\label{HGorderParameters}
    Headgroup and glycerol backbone order parameters of POPS measured in this work compared
    with values for DOPS ($^2$H NMR, 0.1M of NaCl) \cite{browning80} and 
    POPC  ($^{13}$C NMR) \cite{ferreira13} from literature. Signs for PS order parameters
    as measured in this work and signs for PC as measured in Refs \cite{??,ferreira16}.
    %POPG from \cite{borle85} contains 10nM PIPES, DPPG from \cite{wohlgemuth80} contains  10mM PIPES and 100mM NaCl,
    %DPPE from \cite{seelig76}, E.coliPE and E.coliPG are from \cite{gally81}.
  }
\end{figure}
